\documentclass[11pt]{article}

% Other packages for formatting
\usepackage[margin=1in]{geometry}
\usepackage{setspace}
\linespread{1.02}
% \onehalfspacing
\usepackage{fancyhdr}
\usepackage{graphicx}             % For including images
\usepackage{titlesec}             % For customizing section titles

\usepackage{amsmath, physics, amssymb}
\usepackage{parskip}

% Set up the custom page style for the first page
\fancypagestyle{firstpagestyle}{
    \fancyhf{} % Clear all headers and footers
    \fancyhead[L]{\LARGE \textbf{Cover Letter}}
    \fancyhead[R]{\Large \href{https://www.mathisgerdes.com}{\textbf{Mathis Gerdes}}}
    \renewcommand{\headrulewidth}{0pt} % Remove the default header rule
    % \fancyfoot[C]{- {\thepage} -}
}

% Regular page style for the rest of the document
\pagestyle{fancy}
\fancyhf{} % Clear all headers and footers
\fancyhead[L]{\textbf{Cover Letter}} % Regular header on subsequent pages
\fancyhead[R]{\textbf{Mathis Gerdes}}
% \fancyfoot[C]{- {\thepage} -}

\usepackage{xcolor}
\definecolor{royalblue}{rgb}{0.2, 0.3, 0.7} % Adjust the RGB values for your preferred shade of blue
\usepackage[colorlinks=true, linkcolor=royalblue, urlcolor=royalblue, citecolor=royalblue]{hyperref}

% Title information
\title{}
\author{}
\date{}

\begin{document}
\thispagestyle{firstpagestyle}

% Header with right-aligned personal information
\noindent
\begin{minipage}[t]{0.5\textwidth}
\phantom{
\today \\
}
\vspace{0.2cm} \\
Faculty of Mathematics \\
Wilberforce Road \\
Cambridge CB3 0WA \\
United Kingdom \\


\end{minipage}
\begin{minipage}[t]{0.5\textwidth}
\flushright
\today \\
\vspace{0.2cm}
University of Amsterdam \\
Science Park 904 \\
Amsterdam 1098 XH \\
The Netherlands \\
\end{minipage}

\vspace{20pt}

\noindent
\textbf{Dear Members of the Infosys-Cambridge AI Lab Selection Committee,}

I am writing to express my strong interest in the position of Postdoctoral Research Associate in AI and Theoretical Physics within the Infosys-Cambridge AI Lab. Currently, I am a final-year PhD candidate at the University of Amsterdam, supervised by Miranda C. N. Cheng and Christoph Weniger.
My research lies at the intersection of machine learning and theoretical physics, with a focus on leveraging computational techniques to address fundamental challenges in quantum field theory.

The Infosys-Cambridge AI Lab's mission to advance scientific discovery through automation and to develop a mathematical understanding of machine learning aligns closely with my research  and goals. My recent work includes:
\begin{itemize}
    \item \textbf{Lattice Quantum Field Theory:} I advanced continuous normalizing flows for efficient sampling from high-dimensional distributions.
    \item \textbf{Calabi-Yau Metrics:} My contributions include the application of machine learning techniques to approximate Ricci-flat metrics with moduli dependence, going beyond traditional numerical approaches.
    \item \textbf{Physics-Inspired AI:} Recently, I proposed a novel framework for diffusion models informed by insights from the renormalization group, exploring how theoretical physics principles can drive innovations in machine learning.
    \item \textbf{Simulation-Based Inference:} To demonstrate how AI can enable robust analysis of complex astrophysical systems, I implemented a scalable, JAX-accelerated pipelines to infer physical parameters of galactic stellar streams.
\end{itemize}

The Infosys-Cambridge AI Lab's emphasis on interdisciplinary collaboration resonates strongly with my background. My research has consistently bridged domains, and I thrive in collaborative environments that connect physics, AI, and applied mathematics.
For instance, my work on equivariant flows for physics leverages group theory to address computational challenges, and my study of Calabi-Yau metrics integrates differential geometry with modern AI techniques.

Thank you for considering my application. I look forward to the opportunity to contribute to the lab's mission of driving advances in AI and science while collaborating with its vibrant community of researchers.

% \vspace{5pt}
\noindent
\flushright
Sincerely, \\
Mathis Gerdes \\
\href{mailto:m.gerdes@uva.nl}{m.gerdes@uva.nl}

\end{document}
